---
title: "Class 11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sonnets will be discussed

more escape stuff

\\w

## Places to play with Regex
regexone.com
Helps you forget that you are lonly and that life is meaningless - good or friday nights

regexgolf.com

regexr.com - Regex writing site

## Back to the script from yesterday

```{r Class 10 work Hacking10-Regex.R}
# Hacking 10: Regular Expressions
# 2/20/2020

# Stringr Package -------------------------

# install and use stringr for better string splitting
#install.packages("stringr")

library(stringr)

#stringr does this sort of thing:
janewords <- c("These are Jane's words. Some more high-falutin' words. Yup!")
str_split(janewords,boundary("word")) # OMG, I'm so happy
str_split(janewords,boundary("sentence")) # OMG, I'm even happier!
str_trim(unlist(str_split(janewords,boundary("sentence")))) #trim whitespace

# Also has a viewer, which is helpful if you're practicing regex

# From Waber's Essay on Regex Poetry
waber<-c("lip stick","lip sticky","lip stack","lip stacky", "lip stuck", "lip stucky", "lips tick", "lips ticky", "lips tack", "lips tacky", "lips tuck", "lips tucky")

#install.packages("htmlwidgets")
#install.packages("htmltools")
library(htmltools)
library(htmlwidgets)
str_view_all(waber,"lip( s|s )t(i|u)cky?")
# Note, errors and dependencies! What do we do? ...
#STILLBROKEN????

# Scanning in Shakespeare's Sonnets --------------------

# let's scan in the sonnets:
# setwd("~/projects/hack/Shakespeare/sonnets")
# Scan in a text, separate by line breaks
sonnets<-scan("ShakespeareSonnets.txt",what="char",sep="\n")
# Remove the header
sonnets<-sonnets[-(1:7)]
head(sonnets)
tail(sonnets)
sonnets[1:15] # First Sonnet

titleposition<-grep("^[0-9]+",sonnets) # grep is used to find titles in the sonnet text. regex expression is used to isolate the numbers. 
# Alt title: find positions separating sonnets (titles)
length(titleposition) #154 sonnets
length(sonnets) #2309 lines

sonnets<-c(sonnets,"END") #insert an END line for looping
# Add "END" to the end of vector (final line)
length(sonnets) #2310
titleposition<-c(titleposition,2310)

#loop returns three lists (initialized below):
sonnetwords.v<-NULL # a bag for a bag of words
sonnetlines.l<-list() #all the lines in every sonnet in a handy list
sonnetwords.l<-list() #all the words, likewise listed
sonnettables.l<-list() # a frequency table for each sonnet (raw counts)
sonnetrelfreq.l<-list() #relative frequencies

for(i in 1:154){ 
  # create a title for the sonnet by pasting sonnet to a number
  # Adds "Sonnet" and the number i to the title of each sonnet.
  sonnetname<-paste("sonnet",i)
  
  # Find begin and end of each sonnet
  
  # Alt:
  # Find the starting position of each sonnet. Uses the title position plus one because the sonnets start the line after the title.
  # Next line finds the end using a similar method
  start<-titleposition[i]+1
  end<-titleposition[i+1]-1
  
  # Chunk 14 lines of sonnet
  # Alt: puts the text of a sonnet in to the sonnetlines vector.
  sonnetlines<-sonnets[start:end]
  sonnetlines.l[[sonnetname]]<-sonnetlines #create list of sonnets, each list element has 14 sonnet lines in it
  # Paste together 14 lines as blob
  sonnet<-paste(sonnetlines)
  # Split the blob into words (now I have a bag of words for the sonnet)
  sonnetwords<-unlist(str_split(sonnet,boundary("word")))
  # Lower case all the words in the bag of words
  sonnetwords<-tolower(sonnetwords)
  # Remove blanks
  sonnetwords<-sonnetwords[which(sonnetwords!="")]
  sonnetwords.l[[sonnetname]]<-sonnetwords #list of bags of words # puts bag of words into their spot on the list
  
  # Build a big bag of words for all the poems (one bag)
  sonnetwords.v<-c(sonnetwords.v,sonnetwords) #collect all the words in a single vector
  
  # Table the words in "sonnet i"
  sonnet.t<-sort(table(sonnetwords),decreasing=T) 
  sonnettables.l[[sonnetname]]<-sonnet.t #list of frequency tables
  sonnetrelfreq.l[[sonnetname]]<-100*(sonnet.t/sum(sonnet.t)) #list of frequency tables, relative frequencies
}

```


## More script work, but I broke it into pieces

```{R Class 11 }
#a list of lists:
allsonnetlists.l<-list(sonnetlines.l,sonnetwords.l,sonnettables.l,sonnetrelfreq.l)
# str(allsonnetlists.l)
saveRDS(allsonnetlists.l,file="sonnetlists.RDS")

sonnetlines.l[[18]][1]
sonnetwords.l[[18]]

# also create a no-punctuation vector for grepping exercises:
# Punctuation zapping.
sonnets.nopunct<-gsub("[,.:;!?-]","",tolower(sonnets))


```

## Sonnet stuff

nchar counts the number of characters in each word

nchar(sonnetwords.v) <- returns an integer

class

max() finds the largest thing in a vector


sonnets[grep("fire",sonnets)] # lines with "fire" in them
Read from inside out
Greps for fire in the vector of sonnets, this saves the index number. Those numbers are used by sonnets[] to return those lines.
Most of these lines can also be seperated into separate lines if we want.

sapply(sonnetlines.l,str_view_all, pattern="fire") # fails, returns an ugliness 
This fails, trying to apply it to all of the lists

## Sonnets
Sonnets are usually a 14 lines, there are some alternatives
Iambic pentameter
3 quatrains and a concluding couplet

sound shapes are a thing

Italian sonnets are ealier than english

```{R}
# Basic matching and grepping ------------------------

# Longest word in the sonnets?
wordlengths<-nchar(sonnetwords.v)
sonnetwords.v[which(wordlengths==max(wordlengths))] # 14 letter words

class(nchar(sonnetwords.v))

# Find keywords
# Read from inside out
# Greps for fire in the vector of sonnets, this saves the index number. Those numbers are used by sonnets[] to return those lines
sonnets[grep("fire",sonnets)] # lines with "fire" in them
str_view_all(sonnetlines.l[[153]],"fire")

sapply(sonnetlines.l,str_view_all, pattern="fire") # fails, returns an ugliness 

# Count strings 
length(grep("will",sonnetwords.v))
sum(str_count(sonnetwords.v,"will")) # in stringr package

# Count number of vowels in sonnets with str_count

str_count(sonnetwords.v, "a|e|i|o|u")

vowels <- c("a", "e", "i", "o", "u")
word<-"supercalifragilisticexpialidocious"
str_count(word,vowels)

vowelcounts <- vector(mode = "integer", length = 5)
names(vowelcounts)<-vowels
for (i in 1:5) {
  countsbyword<-str_count(sonnetwords.v, vowels[i])
  vowelcounts[i]<-sum(countsbyword) }
vowelcounts # BONUS: can you figure out how to do this with an apply function?

```

## Grepping and sapply
greped and got nothing with the flavor of character


unlist(sapply(sonnetlines.l, grep, pattern = "youth", value = TRUE))
This is an option, but it is not very pretty

purr is another option if you want to get deeper into R.


```{R}
# Playing with grep and sapply over these lists --------------------
# This creates a subection

grep(sonnetwords.l,pattern="fire") #which sonnets are these?
sonnetwords.l[[45]]
sapply(sonnetlines.l,grep,pattern="water",value=T) # but note, 
# empty list elements where there's no match. Is there a more compact way to print?

# A few years ago, Clay Ford (R guru in Brown Library) showed me one way to drop all the empty list elements that are returned
# Two ways, in fact: 1. Base R and using the purr package (I've commented out the second, advanced way)

# example data
test <- list(a = letters[1:3], b = letters[4:6])

# Dummy base R example:
out <- sapply(test, grep, pattern = "a", value = TRUE)
keep <- sapply(out, function(x) length(x)!=0)
out[keep]

#Here's my grepping over the list in base R:
out<-sapply(sonnetlines.l, grep, pattern = "youth", value = TRUE)


keep<-sapply(out, function(x)length(x)!=0)
out[keep]

# # There's another way. This is the "Hadley Wickham way," using purrr package
# # map() is Wickham's version of lapply/sapply
# library(purrr) # for map() and keep(), you will need to install the purrr package
# out2 <- map(test, grep, pattern = "a", value = TRUE)
# keep(out2, function(x)length(x)!=0)
```


## Fancy greping

regex can be just metacharacters

\1 

variations of repetirion

Regex to find poetic structures

(\\w+) is a way to get a word, parenthesis capture it which allows us to reuse it later.

Some of these may crash R.

```{R}

# Tricky grepping ---------------------------

#find ABAB and ABBA patterns? 

# doubled words (Waber: this regex "contains zero literal characters, only meta-characters...")
grep("\\s+(\\w+) \\1\\s+",sonnets.nopunct,value=T)

#ploce? # Does not work right
grep("(\\b[a-z]{3,8}\\b)( \\w+)* \\1(\\w+)+",sonnets.nopunct,value=T) #repeats twice, but filter out the, thee and you, your?
str_view(sonnets.nopunct,"(\\b[a-z]{3,8}\\b)( \\w+)* \\1(\\w+)+",match=T)

# "thyself ... thy ... self" matches
grep("thyself( \\w+)* thy( \\w+)* self", sonnets.nopunct,value=T)
grep(".{0,4}(self|selves)", sonnets.nopunct, value = T)


# ABABs 
# Trying to find parallel structures
grep("(\\w+) (\\w+) (\\w+ )*\\1 \\2",sonnets.nopunct,value=T) # ABAB, simple bigrams: my love, my love
grep("(\\w+) (\\w+) (\\w+ )*\\1 (\\w+ )*\\2",sonnets.nopunct,value=T) # ABABs (more!)
grep("(\\w+) (\\w+ )*(\\w+) (\\w+)* \\1 (\\w+ )*\\3",sonnets.nopunct,value=T) # more ABABs, takes a while

# Compare these in stringr
str_view(sonnets.nopunct,"(\\w+) (\\w+) (\\w+ )*\\1 (\\w+ )*\\2",match=T)
str_view(sonnets.nopunct,"(\\w+) (\\w+ )*(\\w+) (\\w+)* \\1 (\\w+ )*\\3",match=T)

# ABBAs
grep("(\\w+) (\\w+ )*(\\w+) (\\w+)* \\3 (\\w+ )*\\1",sonnets.nopunct,value=T) # ABBA
str_view(sonnets,"(\\w+) (\\w+ )*(\\w+) (\\w+)* \\3 (\\w+ )*\\1",match=T)
str_view(sonnets.nopunct," (\\w+) (\\w+) (\\w+ )*\\2 (\\w+ )*\\1",match=T)

# Lots of ways to write a regex...
grep("(\\w+) (\\w+|\\w+ \\w+|\\w+ \\w+ \\w+) (\\w+) (\\w+|\\w+ \\w|+|\\w+ \\w+ \\w+) \\1 \\3",sonnets.nopunct,value=T)
# [1] "as fast as thou shalt wane so fast thou grow'st" "and chide thy beauty and thy straying youth"    
# [3] "to what you will to you it doth belong"          "thou art as fair in knowledge as in hue"  

#compare
grep("(\\b[a-z]{2,8}\\b ).*( \\b[a-z]{2,8}\\b ).*\\2.*\\1",sonnets.nopunct,value=T)
# [1] "more than that tongue that more hath more expressed"
# [2] "but day by night and night by day oppressed"        
# [3] "or any of these all or all or more"                 
# [4] " the pain be mine but thine shall be the praise"    
# [5] "take all my loves my love yea take them all"        
# [6] "of hand of foot of lip of eye of brow"              
# [7] "none else to me nor i to none alive"                
# [8] "thou art as tyrannous so as thou art"               
# [9] " who art as black as hell as dark as night" 

#compare:
grep("(\\b[a-z]{4} ).*(\\b[a-z]{4} ).*\\1.*\\2",sonnets.nopunct,value=T)
# [1] "as fast as thou shalt wane so fast thou grow'st"   
# [2] "thou dost love her because thou know'st i love her"
# [3] "against that time if ever that time come"          
# [4] " if thou dost seek to have what thou dost hide"  

grep("(\\b[a-z]{2,8}\\b ).*( \\b[a-z]{2,8}\\b ).*\\1.*\\2",sonnets.nopunct,value=T)
# [1] "with sun and moon with earth and sea's rich gems"   
# [2] " lo thus by day my limbs by night my mind"          
# [3] "thou dost love her because thou know'st i love her" 
# [4] "the first my thought the other my desire"           
# [5] "and i am still with them and they with thee"        
# [6] " being had to triumph being lacked to hope"         
# [7] "and all my soul and all my every part"              
# [8] "some glory in their birth some in their skill"      
# [9] "some in their wealth some in their body's force"    
# [10] "some in their hawks and hounds some in their horse" 
# [11] "of hand of foot of lip of eye of brow"              
# [12] "the mountain or the sea the day or night"           
# [13] "that it nor grows with heat nor drowns with showers"
# [14] "but then my friend's heart let my poor heart bail"  
# [15] " who art as black as hell as dark as night"

```

