---
title: "Class 12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Midterm review / Prep
Part I. Fill in the blanks
Part II. Data Structures
Part III. Subsetting
Part IV. Commenting Code

Next week will be statistics.


## How to print out Waldo?

print(here)

```{r }
# Exam Practice (2/21/2019)

# Where's Waldo? -------------------

here<-"Waldo"
there<-"Falstaff"
Ohio<-"Joe"
# Where's Waldo? (That is, what would you write at the R prompt to get "Waldo" to return?)

print(here)
```

# How to get Waldo?
people[[2]]
people[[2]][1]
note the two brackets, one brackets gets the list, we just want Waldo.

class(people[1]) => list
class(people[[1]]) => character

```{r} 
people<-list(there,here,Ohio)
# Now where's Waldo?
people[[2]]
people[[2]][1]

```

## Probably more cruel than the actual exam
Matrix fun
Matrixes are rows then columns

Just keep playing these games. The Waldo questions text understanding of objects.

```{r }

# Put numbers 1 - 6 into a matrix with 3 rows and 2 columns
mymatrix<-matrix(1:6,3,2)

for(i in 1:3) {
  # Loop to 3, i picks out the person in people list
  individ<-people[[i]]
  mymatrix[i,1]<-individ # writes that person into the matrix at row i, column 1
  mymatrix[i,2]<-letters[2*i] # Writes a letter from letters into row i, column 2
  }
# Where's Waldo?
mymatrix[2,1]


# List with vectors, x[[2]] returns a vector
# There will be a question a lot like this on the exam
x<-list("Hal",c("Joe","Waldo"),"Douglas")
# Where's Waldo?
x[[2]][2]

y<-list("Hotspur",x)
# Where's Waldo?
y[[2]][[2]][2]

```

## Dataframes time
factors are a thing, they are wierd

rows are always data.frame objects, columns can be other types such as integer or character

```{R}
# Dataframes --------------------

df<-data.frame(x=1:5,y=10:14,z=letters[1:5],stringsAsFactors = F)
df
class(df[2,])
class(df[1:2,])
class(df[,2:3])
class(df[,2]) # Integer
class(df[2,][2]) 
class(df[2,][[2]]) # bonus, predict what is returned
class(df[,3]) # Character
class(is.na(df[,3][6])) # df gives an NA, is.na() then returns true.

df2<-data.frame(x=1:5,y=10:14,z=15:19,stringsAsFactors = F)
class(df2[2,]) #data.frame
```

## Python????
>>> i = None
>>> type(i)
<class 'NoneType'>

numeric = decimal
integer = integer

```{R}

# Much Ado About Nothing ------------------
# This goes beyond the exam

class("")
class(NULL)
class(NA) # huh...
length(character(0))
length(integer(0))
length(NA) # huh!
length(NULL)
length("")
0+0
0+NA
0+NULL
NA+NA
NA+NULL
NULL+NULL
```

## More classes

Integer vs Numeric, not a big deal for this exam

Vectors are like matricies

lists are like dataframes

https://stackoverflow.com/questions/24052158/why-is-the-class-of-a-vector-the-class-of-the-elements-of-the-vector-and-not-vec

rows then columns - RC Cola




```{R }

# Predict the Class! --------------------
# Predict first, then execute the line to see if you were right.

class("I didn't get an F on my exam because I studied!")
class("F")
class(F)

class(pi)
class(1:100)
class(-1:1)
class(2)
class(2L)
class(-2)

class("LETTERS")

#Remember that strsplit gives lists
class(unlist(strsplit("R is angry and must be propitiated!",split="\\W")))

class(length(letters))

class(class(T))

# rbind makes things into matrixes
class(rbind(x=(1:5),y=10:14))

# dim gives dimensions
class(dim(rbind(x=(1:5),y=10:14)))

class(data.frame(rbind(x=(1:5),y=10:14)))

class(data.frame(rbind(x=1:5,y=10:14))[1,])

class(data.frame(rbind(x=1:5,y=10:14))[,1])

# Data will be turned coalesced into characters
x <- c(1:3)
y <- letters[1:3]
rbind(x,y)

```

## Poetry time

We then described poetry

