---
title: "Class 7"
output:
  html_document: default
  pdf_document: default
---

## Required files
1. ado-actstabled.RDS


## Import 

Lots of people are having trouble getting the RDS objects imported, but it can by done by making the RDS from the script

```{r setup }
knitr::opts_chunk$set(echo = TRUE)

speakerraw.l<-readRDS("ado-speakerrawlist.RDS")
actfreqraw.l <- readRDS("ado-actstabled.RDS")
str(actfreqraw.l)

actfreqraw.l[[1]][1:10]

```

## Matracies and Data frames

Sum adds things together 

lapply first
applies the funtion to the object returning the output as a list
always returns a list

sapply returns a vector
  s stands for simplify
  
there is also just apply

do.call also can take a function and apply to a data structure
Tries to bring all of the data together in some way
applies to the list of the whole

apply vs loops 
very similar, but not the same
R prefers apply
apply is built to work with vectorized data

matracies have 2 demensions
rbind makes matrixes
  RowBind

R does rows then columns




```{r Matracies and Data Frames}

sum(1,3) #Adds things together
mylist <- list(c(1,2,3), c(4,5,6))
sum(mylist[[1]]) # 6

#lapply
lapply(mylist, sum)

#sapply
sapply(mylist, sum)

do.call(sum, mylist)

rbind(1:3, 4:6)
#To make a matrix
mymatrix <- rbind(1:3, 4:6)

#How to get 5 back
mymatrix[2,2]

#Does not work, out of bounds.
# mymatrix[1,4]
# Error in mymatrix[1, 4] : subscript out of bounds

```

## Cbind
cbind - also makes matricies
  column bind

```{r cbind}

mymatrix2 <- cbind(1:3, 4:6)
mymatrix2[2,2]

```

## Bigger matrix time

Missmatched column lengths  repeat the shorter column

```{r bmatrix }

mymatrix <- rbind(5:10, 16:21, 100:105)
mymatrix

apply(mymatrix, 1, nchar)

somenums <- cbind(a=c(3, 2, 4, 5), b = 1:15) # Mismatched column lengths, but it still works
somenums <- cbind(a=c(3, 2, 4, 5), b = 1:4)
somenums
apply(somenums, 1, sum)

```


## lapply

lapply() using "[", the select operator - this is bad


## anewlist
matrix is good enough to auto wrap
to get an item from the matrix coordinates can be used or a list style position can be used.
matrix3[5] goes through the matrix till it hits the 5th item.

lapply(anewlist, "[",,2)
  This is just going to take the second column out
  This gives a warning, but it runs


```{r}
anewlist <- list(matrix(1:9, 3,3),matrix(4:15, 3, 3), cbind(8:10, 8:10))
str(anewlist)

mymatrix3 <- matrix(1:9, 3,3)
mymatrix3
mymatrix3[2,2]
mymatrix3[5]

lapply(anewlist, "[",,2)

```

## English major stuff
It will go act by act and sum the words in each act

lapply will return a list

sapply returns a more compact vector


is.na helps you look for nothings so you can zap it/ replace it with 0



```{r}
lapply(actfreqraw.l, sum)

sapply(actfreqraw.l, sum)

actfreqraw.l[[1]]["nothing"] # Looks for how many times nothing appears in each act

nothings.m <- sapply(actfreqraw.l, "[", c("never", "none", "not", "nothing"))
nothings.m

is.na(nothings.m) # Looks for NAs
which(is.na(nothings.m))

nothings.m[18] <- 0
nothings.m

```

## Looking for character names

actfreqraw.l[[1]]["beatrice"] looks for where beatrices name is used by the characters
it is not used in the first act

NA still infects things, so we have to fix things


```{r}

actfreqraw.l[[1]]["beatrice"]

unlist(lapply(actfreqraw.l, '[', "beatrice"))
# The column name gets infected by NA

ladies.m <- do.call(rbind, lapply(actfreqraw.l, "[", c("beatrice", "hero")))
colnames(ladies.m)
colnames(ladies.m)[1]<-"beatrice"
colnames(ladies.m)

```

## Homework excersize related stuff

this thing is called a term document matrix



NOTE: Document term matrixes are different

```{r}

nothings.m <- sapply(actfreqraw.l, '[', c("never", "none", "not", "nothing"))
nothings.m[18] <- 0


```