---
title: "Class 25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Opening

Final exam, we will review on Tuesday

We are going back into high dimensional space

## Data Wrangling

We are also going to train a support vector machine


## Mrs. Bates, who is this what is this
Pages 252-253

Word vomit

Anything and everything just regurgitated back out

Syntactically free flowing, sporadic bursts

Description is a good place to start when looking at a text

Mrs. Bates as Emma's opposite, something about Mrs. Bates that disturbs Emma

This is direct discourse
there are things that Emma can't see and therefore we can't see, but they become clearer when we read it again
* Frank Churchill's eagerness to help Emma down from her carriage

Jane Austen style and character style

"Do not MIMIC her"

## Classification
Supervised learning

Show computer examples and the computer creates a model

Model works to find dividing line between two different groups of things

What about ambiguous passages - they are really close or even over the line

What about mistakes - as a fan writing fiction, you want your work to be the mistake.

## Trained model defines a hyperplane and margins



```{r Hacking25-ClassifyingSVM.R}
# Hacking, Class 22
# SVM Classification
# April 22, 2020

# Oh man, this is complicated stuff, but read here for a cookbook approach:
# https://www.csie.ntu.edu.tw/~cjlin/papers/guide/guide.pdf

# And more here: https://www.csie.ntu.edu.tw/~cjlin/libsvm/

# install.packages("e1071")
library("e1071") # svm package is in here
source('Rtools.R') # our tools: using getcorpus(). I'll paste in the same function below.
# setwd("~/projects/hack/austen")

# More bits: from http://pemberley.com/?page_id=5270
bitsofivory<-getcorpus("morebits") # 22 bits, all emma rewrites
mean(sapply(bitsofivory,length)) # Average length is about 5689 (we'll need this below)

# scan and clean emma
emma.text<-scan("austen-emma.txt",what="char",sep="\n")
emma<-unlist(strsplit(tolower(emma.text),"\\W"))
emma<-emma[which(emma!="")]

# Open the file for consulting (use the file menu)?

# chunk the novel into segments that are 5689 words long 
x<-seq_along(emma)
bitsofemma<-split(emma,ceiling(x/5758)) #returns 29 chunks, last is only 896 words
names(bitsofemma)<-paste0("emma",seq_along(bitsofemma)) #name list elements emma1, emma2, etc.

str(bitsofemma)

# combine all the bits
bits.l<-c(bitsofemma,bitsofivory)
str(bits.l)

# Our research Q: Can we tell bits of austen from bits of ivory?
# I'm following Jockers data munging in Chapters 11 and 12, adapting it to my texts. 
```


```{r} 
# first calculate relative word frequencies
bitsofemma.freq.l<-lapply(bitsofemma,table)
bitsofivory.freq.l<-lapply(bitsofivory,table)
bitsofemma.relfreq.l<-lapply(bitsofemma.freq.l,prop.table)
names(bitsofemma.relfreq.l)<-paste0("emma",seq_along(bitsofemma.relfreq.l))
bitsofivory.relfreq.l<-lapply(bitsofivory.freq.l,prop.table)

# a new function: written by Clay Ford to solve my Jockers' problems (thank you, Clay!).
toDF <- function(x,j){
  data.frame(ID = j,
  Word = unlist(dimnames(x)),
  Freq = as.data.frame(x)$Freq,
  stringsAsFactors = FALSE)
}

# use that function, applying it to the lists of frequencies.
emma.freqs.l <- mapply(toDF,
                         x = bitsofemma.relfreq.l,
                         j = names(bitsofemma.relfreq.l),
                         SIMPLIFY = FALSE) # keep it as a list

ivory.freqs.l <- mapply(toDF,
                         x = bitsofivory.relfreq.l,
                         j = names(bitsofivory.relfreq.l),
                         SIMPLIFY = FALSE)

# check on the output
# "about" is a fan word
emma.freqs.l[[2]][1:10,]
ivory.freqs.l[[2]][1:10,]

#str(emma.freqs.l) # a list: each dataframe element with bit id, words, and frequencies

#bind list of dataframes into one dataframe
emma.freqs.df<-do.call(rbind.data.frame,emma.freqs.l)
row.names(emma.freqs.df)<-NULL
ivory.freqs.df<-do.call(rbind,ivory.freqs.l)
row.names(ivory.freqs.df)<-NULL

# Prepare dataframes for classification
allthebits.freqs.df<-rbind(emma.freqs.df,ivory.freqs.df) # one big dataframe
dim(allthebits.freqs.df) # long rectangular data: 57586 rows, 3 columns

# Cross tabulation: go from long-form to wide-form table
result.t<-xtabs(Freq ~ ID+Word,data=allthebits.freqs.df)
# 51 Chunks of text
# 9716 unique terms
# Getting into a document term matix
dim(result.t) #51 rows, 9716 columns

# put labels in the dataframe
final.df<-as.data.frame.matrix(result.t) 
final.df[1:10,65:80] # looks like a proper DTM, good!
chunks<-rownames(final.df) # use rownames to create two labeled classes
chunks<-gsub(".+txt","fan",chunks) # fans label
chunks[5:33]<-"austen" # austen label
final.df<-cbind(final.df,chunks) # now can pull out labels with final.df$chunks
final.df[1:51,c("emma","knightley","marriage","chunks")] # peek at it, looking at just 3 columns, compare frequencies of character mentions in fan fiction and in "Emma"
dim(final.df) # 51 9717 -- That's big!
#View(final.df[, 100:150])

# reduce the size of the dataframe
freqmeans.v<-colMeans(final.df[,1:9716]) # find average frequency, across corpus, for each word
greaterthans.v<-which(freqmeans.v>=.001) # choose words/columns greater than .1%

smaller.df<-final.df[,c(names(greaterthans.v),names(final.df[9717]))]
dim(smaller.df) # 51 144 -- that's better: 51 docs and 144 features.
View(smaller.df)

```

```{r} 

###########################################################################

# THE MAIN EVENT...

# Note, you can run the following repeatedly and will get different results.
#RUN AND RERUN FROM HERE DOWN >>>>>>>>>>>>>>

# choose, randomly, half the texts to train on
randomrows<-sample(nrow(smaller.df),nrow(smaller.df)/2) # sample of half the rows, choosing half the chunks to train on.
train<-smaller.df[randomrows,] # train on these texts
test<-smaller.df[-randomrows,] # test on these
dim(train)
dim(test)

# train the model
model.svm<-svm(train[1:143],train$chunks)
summary(model.svm)

# apply the model to the training data
pred.svm<-predict(model.svm,train[1:143])
as.data.frame(pred.svm)
summary(pred.svm)


# Breaking???
# table(pred.svm,train$chunks) # confusion matrix

# apply the model to the test data
testing<-predict(model.svm,test[1:143], decision.values=T)
predict(model.svm,test[1:143], decision.values=T) # see separation scores for each bit
as.data.frame(testing) # see all the labeling the classifier did. Note, emma29 may be miscategorized as fan fic
summary(testing)
table(testing,test[,144]) # confusion matrix. Watch for false negatives (classifier predicted a fan bit, got an Austen bit!)

# Sometimes when I run this, I get emma29 confused as fan fiction (a shorter bit)
# first time, emma29 mistaken for fan fiction (a false negative)
# second time, all were right
# third time, I got emma29 again
# fourth time, emma29 again

# WHY? is the final chunk mistaken? —Something to think about.
# DISCUSS: because a happy ending, fans reading for marriage, Knightley and Emma? 

# Write the prediction factors into a dataframe so that we can sort through texts
predictionfactors<-predict(model.svm,test[1:143], decision.values=T)
# decisval.m<-attr(predictionfactors,"decision.values") # pull out a matrix?
# Actually, this is better:
decisval.df <- data.frame(text=names(predictionfactors),
                          score=as.numeric(attr(predictionfactors,"decision.values")),
                          stringsAsFactors = F)
decisval.df<-decisval.df[order(decisval.df$score),] # order the texts
View(decisval.df)

# First run:
# See that laraine-knightsinblacksatinwaistcoats.txt is closest to the hyperplane. Read it!
# See that ulrike-alovemostunwanted1.txt is the farthest from the hyperplane. Read it, if you dare!
# Although, not, it's also 3 times as long as the average bit (it's 15,467 words, average is 5689)
# I should probably do this again, but scale better? 

# It appears (maybe) fan fic with lots of dialogue (or letter writing in it) is easy to classify as *not* Austen. Too many first-person pronouns?

# Literary critics will want to look at the words scored...
model.svm$SV # look at this thing
dim(model.svm$SV) # 24 by 143
w = t(model.svm$coefs) %*% model.svm$SV # get feature weights
weights.df<-sort(as.data.frame(w)) 
View(weights.df) # top fan words
View(sort(weights.df,decreasing=T)) #top Austen words
 
##################################################################

# from earlier script (Looking at Pride and Prejudice fan fiction:
# get bits of ivory corpus
bitsofivory<-getcorpus("bitsofivory") #84 bits
mean(sapply(bitsofivory,length)) # Average length is 6684.655 (we'll need this below)

# scan and clean pride and prejudice
pride<-scan("austen-pride.txt",what="char",sep="\n")
pride<-gsub("_","",pride)
pride<-unlist(strsplit(tolower(pride),"\\W"))
pride<-pride[which(pride!="")]

# chunk the novel into segments that are 6685 words long
x<-seq_along(pride)
bitsofausten<-split(pride,ceiling(x/6685)) #returns 19 chunks of 6685, on of 2540

# Interesting false positive when working with Pride and Prejudice.
# Note: as.data.frame(testing) # see all the labeling the classifier did. Note, pride4 is miscategorized as fan fic

################################################

# In case you can't find your RTools, or don't have a getcorpus function in it:

# This function extracts files from a directory (dir), atomizes as words, 
# and puts them in a list
getcorpus<-function(dir,type=".txt"){
  library(stringr)
  curr.folder<-getwd()
  setwd(dir)
  corpus<-list()
  files<-list.files(pattern=type)
  for(i in 1:length(files)){
    text<-scan(files[i],what="char",sep="\n")
    text<-paste(text,collapse=" ")
    lowertext<-tolower(text)
    text.words<-unlist(str_split(lowertext,boundary("word")))
    text.words<-text.words[which(text.words!="")]
    corpus[[files[i]]]<-text.words
  }
  setwd(curr.folder)
  return(corpus)
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
