---
title: "Class 13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Test review


```{r test questions}
# 2.
charnames <- c("Waldo", "Dogberry", "Shakespeare", "Claudio", "Waldo")
charnames.t <- sort(table(charnames), decreasing = T)
names(charnames.t)[[1]]

# 3.
flist <- c(letters[1:3], "Waldo", 1:10)
glist <- list(letters[4:6], flist)
glist[[2]][[4]]

# 4.
df <- data.frame(cbind(1:3, c("Larry", "Moe", "Waldo"), LETTERS[24:26]), stringsAsFactors = F)
slice.df <- df[3,]

# 5.
df <- data.frame(1:3, c("Larry", "Moe", "Waldo"), LETTERS[24:26], stringsAsFactors = F)
slice <- df[,2]
slice[[3]]

```

## Back to regularly scheduled program

Tf-idf

hapax legomena - words that only appear once.

Term Frequency (tf)
Term frequency is the number of times a term t occurs in document d

Document Frequency (df)
Document frequency is the number of number of times a document d with a term t in it occurs in a corpus of N documents.

Term Frequency, Inverse Document Frequency (tf.idf)
tf.idf = tf * idf
tf.idf = tf * log(N/df)

Terms that show up in all documents have a high document frequency score, but a low inverse document frequency score.


## Sonnet 18
Summer seems important, but it also shows up in a lot of other sonnets.
summer -> tf-idf is 7.4



```{r }

sonnetsrawtables.l <- readRDS("sonnetlists.RDS")[[3]]

tf <- sapply(sonnetsrawtables.l, '[', "eye")
tf[which(is.na(tf))] <- 0 # This replaces nA with a 0 value.
sum(tf) # Term frequency across all sonnets
tf[18] # tf for sonnet 18, it's 1

tf!=0

sum(tf)
sum(tf!=0) # df

#tf-idf score for eye
1*log(154/27)


tf <- sapply(sonnetsrawtables.l, '[', "the")
sonnet <- 18
tf[which(is.na(tf))] <- 0 # This replaces nA with a 0 value.
sum(tf) # Term frequency across all sonnets
tf[sonnet] # tf for sonnet 18, it's 1

tf!=0

sum(tf)
sum(tf!=0) # df

#tf-idf score for eye
tf[sonnet]*log(154/sum(tf!=0))


tf <- sapply(sonnetsrawtables.l, '[', "sweet")
sonnet <- 54
tf[which(is.na(tf))] <- 0 # This replaces nA with a 0 value.
tf[sonnet] # tf for sonnet 18, it's 1
sum(tf) # "Sweet" appears 56 times in sonnets
sum(tf!=0) # df

#tf-idf score for eye
tf[sonnet]*log(length(tf)/sum(tf!=0))


# Time to get more complicated

# now we are thinking with applys.

terms <- c("sweet", "old", "youth", "hate" )

tf.m <- sapply(sonnetsrawtables.l, '[', terms)
rownames(tf.m) <- terms
tf.m[which(is.na(tf.m))] <- 0
View(tf.m)

df.m <- rowSums(tf.m != 0)

```


