---
title: "Class 8"
output:
  html_document: default
  pdf_document: default
---

## Required files
1. muchado-edgelist.txt
1. Hacking8-Network.R



## Today's Commands
install.packages()
library() -> require can be used, not sure what the exact differences are
{igraph} A library for network analysis: igraph.org
subset.data.frame()
graph.data.frame{igraph}


## Beginning of class


Thursday will be focused on Much ado

We talk about graphs and stuff

Can catch flavor of an author from function words - other thigns

Seven Bridges of Konigsberg
- First graph theory problem

Finding Paul Revere using only metadata

Adjacency Matrix and resulting graph
-Only need 1 half unless it is a directed graph

We had issues with the AV system

Discussed onion article titled: Professor having trouble with audio visual
https://www.theonion.com/ask-a-college-professor-having-trouble-with-the-audiovi-1819584653

We then watch the Watch and Constable Dogberry sceen - the one with the Sexton


## Script explained
install.packages("igraph") only needs to be done once
-Sometimes dependencies will make this more difficult

library("igraph") then attaches igraph to the session

we then import muchado-edgelist.txt as a dataframe
dataframe allows for multiple data types.
A matrix would not allow for different data types.

ado.df[1<=ado.df\$act.sc & ado.df\$act.sc<2,]
Give me everything in scene 1 and 2


```{r Hacking 8 Networks}
# Social Networks in Literature
# February 11, 2020

# install.packages("igraph") # Only needs to be done once.
library("igraph") # Attaches igraph to the session
#setwd("~/projects/hack/igraph")

# Reading tables ------------
# Read in information about dialogue in Much Ado...
# I've created a representation of who speaks to whom in each scene, marking exchanges only once per scene.
# That is if Beatrice and Benedick go back and forth for several lines, I record only:
# BEAT\tBEN and BEN\tBEA
# soliloquy and asides and oaths are marked with an "x"
ado.df<-read.table('muchado-edgelist.txt', header=T) # strings come in as factors, compare read.table('muchado-edgelist.txt', header=T, stringsAsFactors=F)
View(ado.df) # a data-frame (mixes data types)! 
class(ado.df[,1]) == class(ado.df[,3]) 

# Subsetting (Tricks) -------- 
ado.df$char1 # factors (categorical variables)
as.character(ado.df$char1) # as a character vector
class(ado.df$act.sc)
ado.df$act.sc
which(ado.df$act.sc==4.2) # Dogberry and "malefactors" before the Sexton
ado.df[which(ado.df$act.sc==4.2),] # same subsetting, more verbose
grep("1.",ado.df$act.sc)
ado.df[grep("1.",ado.df$act.sc),]
subset.data.frame(ado.df, char1=="DOGB")
1<=ado.df$act.sc & ado.df$act.sc<2
ado.df[1<=ado.df$act.sc & ado.df$act.sc<2,]
act1sc3<-ado.df[which(ado.df$act.sc==1.3),]

# Try out the igraph library -----------
g <- graph.data.frame(ado.df)

str(g)

V(g) # vertices, there are 23
E(g) # edges, there are 246

#get all the edges
get.edgelist(g)
class(get.edgelist(g)) #it's a matrix

# plots
plot(g) # The whole play: whoa, that's a spaghetti monster!
plot(g, layout=layout.circle)
plot(g, layout=layout.fruchterman.reingold)
plot(g, layout=layout.graphopt)
plot(g, layout=layout.kamada.kawai)

# prettify that plot? 
# prettier plots: http://igraph.org/r/doc/plot.common.html
par(mai=c(0,0,1,0))
plot(g,    		#the graph to be plotted
     layout=layout.kamada.kawai,	# the layout method. see the igraph documentation for details
     main='"Much Ado - Addressed Speech',	#specifies the title
     vertex.size=4,
     edge.arrow.size=.3,
     vertex.label.dist=0.5,			#puts the name labels slightly off the dots
     vertex.frame.color='blue', 		#the color of the border of the dots 
     vertex.label.color='black',		#the color of the name labels
     vertex.label.font=2,			#the font of the name labels
     vertex.label=V(g)$name,		#specifies the labels of the vertices.
     vertex.label.cex=.5			#specifies the size of the font of the labels
)

# Simplify 
g2<-graph.data.frame(unique(ado.df[,1:2]))
par(mai=c(0,0,1,0))
plot(g2, layout=layout.kamada.kawai, main='"Much Ado - Unique Addressed Speech',
     vertex.size=4, edge.arrow.size=.3, vertex.label.dist=0.5,
     vertex.frame.color='blue', vertex.label.color='black',	
     vertex.label.font=2, vertex.label=V(g)$name,vertex.label.cex=.5
)

# plot and save as a PDF
pdf("ado-who2whom.pdf")
plot(g2)
dev.off()

# Simplify: Pull out a scene?
ado.df[,3] #ask for a column by number
ado.df$act.sc #ask for a column by name
ado.df[which(ado.df$act.sc == 4.2),] # subset rows
act4sc2<-ado.df[which(ado.df$act.sc == 4.2),]
g4<-graph.data.frame(act4sc2)
plot(g4)

V(g) # 23/23 vertices, named:
# + 23/23 vertices, named, from 4e6264d:
#  [1] LEON  MESS  BEAT  HERO  PEDRO BEN   JOHN  CLAUD ANTON CONR 
# [11] BORA  BALTH MARG  URSU  BOY   DOGB  VERG  WATCH SEAC  FRIAR
# [21] SEXT  x     LORD 

# Adjacency
neighbors(g, V(g)[3], mode = "out") #who does Beatrice talk to? (18 of 23)
neighbors(g, V(g)[6], mode = "out") #who does Benedick talk to? (28 of 23)
neighbors(g, V(g)[5], mode = "out") #Don Pedro? (33/23)
neighbors(g, V(g)[1], mode = "out") #who does Leonato talk to? (40 of 23)
neighbors(g, V(g)[3], mode = "in") #who talks to Leonato (17 of 23)
neighbors(g, V(g)[4], mode = "out") #who does Hero talk to? (13 of 23)
neighbors(g, V(g)[7], mode = "out") #Don John? (33/23)

# Incidence (edges that share a vertex)
incident(g,V(g)[3], mode="all")
incident(g,V(g)[3], mode="out")
incident(g,V(g)[3], mode="in")

# indegree and outdegree socialty (quiz and sort these)
degree(g, mode="in") # spoken to
degree(g,mode="out") # speaks 

# paths
View(distances(g))
View(shortest.paths(g, mode='in'))
View(shortest.paths(g, mode='out'))
# all_shortest_paths(g,from="HERO") #super long text
all_shortest_paths(g2,from="DOGB",to="JOHN")

# betweenness: measures the number of shortest paths going through a specific vertex
betweenness(g)

# eigenvector centrality: gives greater weight to a node the more it is connected to other highly connected nodes
g_undirected <- as.undirected(g, mode='collapse')
evg_social<-evcent(g_undirected)
eigen_social<-evg_social$vector
sort(eigen_social,decreasing=T)

# closeness: "closeness centrality of a vertex is defined by the inverse of the average length of the shortest paths to/from all the other vertices in the graph"
closeness(g, mode='in') # WTF? -- Why warning?
components(g) # also try, component_distribution(g)
is_connected(g)
components(g,mode="strong") # AH! In fact, 3 graphs!
is_connected(g,mode="strong") #My "x" and my "LORD don't talk back!
g2<-induced_subgraph(g, which(components(g,mode="strong")$membership==1)) # Fix it
closeness(g2)

# This should also work?
# which(ado.df$char2!="x")
# ado2.df<-ado.df[which(ado.df$char2!="x"),]
# ado2.df<-ado2.df[which(ado.df$char2!="LORD"),]
# g3<-graph.data.frame(ado.df2)

# Some more subsetting, tips and tricks
act1sc3<-ado.df[which(ado.df$act.sc==1.3),]
act1scene3.g<-graph.data.frame(act1sc3)
str(act1scene3.g)
act1scene3.g[1]
neighbors(act1scene3.g,V(act1scene3.g)[2],mode="out")
V(act1scene3.g)
degree(g)

##############
# START HERE? -------------
# messing about also with this: https://rstudio-pubs-static.s3.amazonaws.com/74248_3bd99f966ed94a91b36d39d8f21e3dc3.html

g<-graph.empty(n=10, directed=TRUE)
plot(g)

g<-graph.full(n=10, directed = FALSE, loops = FALSE)
plot(g)
g<-graph.star(n=10, mode="out")
plot(g)
g<-graph.star(n=10, mode="in")
plot(g)
g<-graph.ring(n=10)
plot(g)

edges <- c(1,2, 3,2, 2,4)
g<-graph(edges, n=max(edges), directed=TRUE)
plot(g)

vcount(g) #count vertices
ecount(g) #count edges

neighbors(g, V(g)[1], mode = 1)
incident(g,V(g)[2], mode=c("all", "out", "in", "total"))
is.directed(g)
are.connected(g, V(g)[1], V(g)[3])
get.edgelist(g)



############################################################
############################################################

# From last time I taught this class: 
# Jane Austen's Emma

# read in information about Emma dialogue
emma.df<-read.table('emma-graph.txt', header=T) # Volume 1, Who speaks to whom?
g <- graph.data.frame(emma.df)

V(g) # vertices, there are 14
E(g) # edges, there are 90

# plots
plot(g)
plot(g, layout=layout.circle)
plot(g, layout=layout.fruchterman.reingold)
plot(g, layout=layout.graphopt)
plot(g, layout=layout.kamada.kawai)

# prettify that plot? 
# prettier plots: http://igraph.org/r/doc/plot.common.html
par(mai=c(0,0,1,0))
plot(g,    		#the graph to be plotted
     layout=layout.kamada.kawai,	# the layout method. see the igraph documentation for details
     main='"Emma" - Addressed Speech',	#specifies the title
     vertex.size=4,
     edge.arrow.size=.3,
     vertex.label.dist=0.5,			#puts the name labels slightly off the dots
     vertex.frame.color='blue', 		#the color of the border of the dots 
     vertex.label.color='black',		#the color of the name labels
     vertex.label.font=2,			#the font of the name labels
     vertex.label=V(g)$name,		#specifies the labels of the vertices.
     vertex.label.cex=.5			#specifies the size of the font of the labels
)

# plot and save as a PDF
pdf("emmagraph.pdf")
plot(g)
dev.off()

V(g) # 14/14 vertices, named: MrWoodhouse, Emma, MrKnightley, Harriet, MrElton, MissNash, Isabella    
    # MrJKnightley, MrWeston ,MrsWeston, Frank, x, MrsBates, MissBates

neighbors(g, V(g)[2], mode = 1) #who does Emma talk to? 
neighbors(g, V(g)[3], mode = 1) #who does Knightley talk to? 

incident(g,V(g)[2], mode=c("all", "out", "in", "total"))
get.edgelist(g)

# indegree and outdegree socialty
degree(g, mode="in") # spoken to
degree(g,mode="out") # speaks to 

# closeness
closeness(g, mode='in')
closeness(g, mode='out')

# betweenness: measures the number of shortest paths going through a specific vertex
betweenness(g)

# eigenvector centrality: gives greater weight to a node the more it is connected to other highly connected nodes
g_undirected <- as.undirected(g, mode='collapse')
evg_social<-evcent(g_undirected)
eigen_social<-evg_social$vector

# paths
View(distances(g))
View(shortest.paths(g, mode='in'))
shortest.paths(g, mode='out')


###############################################
# messing about also with this: https://rstudio-pubs-static.s3.amazonaws.com/74248_3bd99f966ed94a91b36d39d8f21e3dc3.html

g<-graph.empty(n=10, directed=TRUE)
plot(g)

g<-graph.full(n=10, directed = FALSE, loops = FALSE)
g<-graph.star(n=10, mode="out")
g<-graph.star(n=10, mode="in")
g<-graph.ring(n=10)

edges <- c(1,2, 3,2, 2,4)
g<-graph(edges, n=max(edges), directed=TRUE)

vcount(g) #count vertices
ecount(g) #count edges

neighbors(g, V(g)[1], mode = 1)
incident(g,V(g)[2], mode=c("all", "out", "in", "total"))
is.directed(g)
are.connected(g, V(g)[1], V(g)[3])
get.edgelist(g)



```

## title

```{r title}



```

## title


```{r  title2}



```

