---
title: "Class 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Book day

Today was a book day.

```{r}


```

# betweenness: measures the number of shortest paths going through a specific vertex
betweenness(g)

# eigenvector centrality: gives greater weight to a node the more it is connected to other highly connected nodes
g_undirected <- as.undirected(g, mode='collapse')

evg_social<-evcent(g_undirected)

eigen_social<-evg_social$vector

sort(eigen_social,decreasing=T)

# closeness: "closeness centrality of a vertex is defined by the inverse of the average length of the shortest paths to/from all the other vertices in the graph"
closeness(g, mode='in') # WTF? -- Why warning?

components(g) # also try, component_distribution(g)

is_connected(g)

components(g,mode="strong") # AH! In fact, 3 graphs!

is_connected(g,mode="strong") #My "x" and my "LORD don't talk back!

g2<-induced_subgraph(g, which(components(g,mode="strong")$membership==1)) # Fix it

closeness(g2)


## Movie time
Joss Weadom and the Branough(sp?)

Benedick overhearing II.iii.87ff

Beatrice overhears III.i.18ff


```{r}

```