---
title: "Class 18"
output: html_document
---

```{r setup, include=FALSE}
```

## R Diving right in
We only need to remove the first 4 lines, not 6

```{r }
songs <- scan("blake-songs-all-clean.txt", "char", sep ="\n")
songs <- songs[-(1:4)] #Negative ask for everything but the first 4 lines
library(stringr)
songs.words <- unlist(str_split(songs, boundary("word")))

# This works as long as we are in the middle of the vector
colloc1<-function(target.kw,text.words,LRspan){
  # get the keyword positions
  target.index<-which(tolower(text.words)==tolower(target.kw))
  collocates<-NULL
  for(i in 1:length(target.index)){
    #collect the window's collocates
    LHS<-text.words[(target.index[i]-LRspan):target.index[i]]  #Grab to the left
    RHS<-text.words[target.index[i]:(target.index[i]+LRspan)] #Grab to the right
    collocates<-c(collocates,LHS[-length(LHS)],RHS[-1])   #Add to previously collected collocates  
  }
  # Return them
  return(table(tolower(collocates)))
}

#colloc1("piping", songs.words, 4)

```

## If and else statements

if(a > 5){print("a is too big")}else{print("OK!")}

It is a kind of switching logic

We are going to use this logic prevent collocator from going over

```{r if and else statements}

a <- 3
if(a > 5){print("a is too big")}else{print("OK!")}

target.kw <- "Introduction"
text.words <- songs.words
which(tolower(text.words) == tolower(target.kw))

(3-5) < 1
(5-4) < 1


# This works as long as we are in the middle of the vector
colloc1<-function(target.kw,text.words,LRspan){
  # get the keyword positions
  target.index<-which(tolower(text.words)==tolower(target.kw))
  collocates<-NULL
  for(i in 1:length(target.index)){
    if(target.index[i]-LRspan < 1){
      LHS<-text.words[1:target.index[i]]
    } else{
      LHS<-text.words[(target.index[i]-LRspan):target.index[i]]
    }
    
    if(target.index[i]+LRspan > length(text.words)){
      RHS<-text.words[target.index[i]:length(text.words)] 
    } else{
      RHS<-text.words[target.index[i]:(target.index[i]+LRspan)] 
    }
    
    #collect the window's collocates
    RHS<-text.words[target.index[i]:(target.index[i]+LRspan)] #Grab to the right
    collocates<-c(collocates,LHS[-length(LHS)],RHS[-1])   #Add to previously collected collocates  
  }
  # Return them
  return(table(tolower(collocates)))
}

colloc1("piping", songs.words, 4)


# How to bring in RTools

source("RTools.R")

```

## Now on to the script from the other day

KWIC

Sometimes words show up in their own company.

```{R Hacking18-KWIC.R}
# Hacking for Humanists
# March 31, 2020

# Collocation: "You shall know a word by the company it keeps!"

songs<-scan("blake-songs-all-clean.txt",what="char",sep="\n")
songs.words<-unlist(strsplit(songs,"\\s")) # when concordancing, we may want to keep punctuation

#better?
library("stringr")
songs.words<-unlist(str_split(songs,boundary("word")))

# Building a Collocator (this one breaks)
colloc1<-function(target.kw,text.words,LRspan){
  target.index<-which(tolower(text.words)==tolower(target.kw))
  cat("The word",target.kw,"appears",length(target.index),"times\n")
  collocates<-NULL
  for(i in 1:length(target.index)){
    LHS<-text.words[(target.index[i]-LRspan):target.index[i]]
    RHS<-text.words[target.index[i]:(target.index[i]+LRspan)]
    collocates<-c(collocates,LHS[-length(LHS)],RHS[-1])
  }
  return(table(tolower(collocates)))
}

# A collocator that doesn't break (I think!)
colloc<-function(target.kw,text.words,LRspan){
  target.index<-which(tolower(text.words)==tolower(target.kw))
  if(length(target.index) == 0){
    print("No match!")
  }
  
  cat("The word",target.kw,"appears",length(target.index),"times\n")
  collocates<-NULL
  for(i in 1:length(target.index)){
    if((target.index[i]-LRspan)<1){
      LHS<-text.words[1:target.index[i]]
    }else{
      LHS<-text.words[(target.index[i]-LRspan):target.index[i]]
    }
    if((target.index[i]+LRspan)>=length(text.words)){
      RHS<-text.words[target.index[i]:length(text.words)]
    }else{
      RHS<-text.words[target.index[i]:(target.index[i]+LRspan)]
    }
    collocates<-c(collocates,LHS[-length(LHS)],RHS[-1])
  }
  return(table(tolower(collocates)))
}

###########################################################

# Kwicking (for Thursday)

# set for testing
target.kw="innocence"
text.words=songs.words

# Mark Algee-Hewitt's KWICker (tweaked and commented by Brad)
kwic<-function(target.kw, text.words, LRspan){
  #find places in text.words where target.kw can be found
  #target.index<-which(text.words==target.kw) #control for casing below
  target.index<-grep(tolower(target.kw),tolower(text.words))
  if(length(target.index)==0){return("No match")}
  #create empty variable
  target.kwics<-NULL
  #loop that repeats for all incidents of keyword in text
  for (i in 1:length(target.index)){
    #check to make sure, we're in bounds
    if (target.index[i]-LRspan<1){
      start<-1
    } else {
      start<-target.index[i]-LRspan
    }
    if (target.index[i]+LRspan>length(text.words)){
      end<-length(text.words)
    }else{
      end<-target.index[i]+LRspan
    }
    #grab chunk
    text.chunk<-text.words[start:end]
    #collapse chunk to a single variable, separated by spaces
    text.chunk<-paste(text.chunk, collapse=" ")
    #add chunk to vector of chunks
    target.kwics<-c(target.kwics, text.chunk)
  }
  #return complete vector of text chunks
  return(target.kwics)
}

kwic("god", songs.words, 7)

kwic("man", songs.words, 7)
```

## Some ways to read in lots of files

Can read in directories at a time

```{r }

#####################################################

# A final function: get a corpus into R

blakedir<-"./songstitled"
filenames<-dir(blakedir)

filepath<-"./songstitled"
# Quickly list files in a directory:
files.v<-dir(filepath,"\\.txt$") # Jockers' way
files.v<-list.files(filepath) # Why not this way?

# Or use a built-in function?
list.files(pattern='.txt')
list.files(pattern='.txt',recursive=T)
dir(blakedir)

# Write a function?
showfiles<-function(filenames){
  for(i in 1:length(filenames)){
    cat(i,filenames[i],"\n",sep=" ")
  }
}

# Convert files into bags of words:
# Jockers' Way

# Takes in an entire directory

makecorpus<-function(files,directory){
  text.words.l<-list()
  for(i in 1:length(files)){
    text<-scan(paste(directory,files[i],sep="/"),what="char",sep="\n")
    text<-paste(text,collapse=" ")
    lowertext<-tolower(text)
    text.words<-unlist(strsplit(lowertext,"\\W"))
    text.words<-text.words[which(text.words!="")]
    text.words.l[[files[i]]]<-text.words
  }
  return(text.words.l)
}

# Brad's way (Again, borrowing from Mark Algee-Hewitt)
# Takes in an entire directory
getcorpus<-function(dir,type=".txt"){
  curr.folder<-getwd()
  setwd(dir)
  corpus<-list()
  files<-list.files(pattern=type)
  for(i in 1:length(files)){
    text<-scan(files[i],what="char",sep="\n")
    text<-paste(text,collapse=" ")
    lowertext<-tolower(text)
    text.words<-unlist(strsplit(lowertext,"\\W"))
    text.words<-text.words[which(text.words!="")]
    corpus[[files[i]]]<-text.words
  }
  setwd(curr.folder)
  return(corpus)
}

getcorpus("songstitled")

# Add these three functions to your myRTools.R file:
# colloc()
# kwickr()
# getcorpus()

# Note, you might improve them as we go by adding the following:
# if(length(target.index)==0){
# warning("No match")}



```